---
title: "hw1"
author: "Rao Yuner"
format:
  html:
    self-contained: true
    code-fold: true
engine: knitr
---

```{r}
#problem 1(a) momentum_walk(n,p)
momentum_walk <- function(n, p) {
  if (length(n) != 1 || !is.numeric(n) || n < 0 || n != floor(n)) {
    stop("n must be a non-negative integer")
  }
  if (length(p) != 1 || !is.numeric(p) || p < 0 || p > 1) {
    stop("p must be between 0 and 1")
  }
  directions <- matrix(
    c(1, 0,
      -1, 0,
      0, 1,
      0, -1),
    ncol = 2,
    byrow = TRUE
  )
  position <- c(0, 0)
  current_direction <- sample(1:4, 1)
  position <- position + directions[current_direction, ]
  if (n >= 2) {
    for (i in 2:n) {
      if (runif(1) < p) {
      } else {
        current_direction <- sample(1:4, 1)
      }
      position <- position + directions[current_direction, ]
    }
  }
  return(position)
}


#set example
momentum_walk(10, 0.6)
```

### Explanation problem1(a)

if we run the same parameters for several times, the outputs are different.

```{r}
#problem1 partb
#version1
momentum_walk1 <- function(n, p) {
  
  if (length(n) != 1 || !is.numeric(n) || n < 0 || n != floor(n)) {
    stop("n must be a non-negative integer")
  }
  
  if (length(p) != 1 || !is.numeric(p) || p < 0 || p > 1) {
    stop("p must be between 0 and 1")
  }
  
  directions <- matrix(
    c(1, 0,
      -1, 0,
      0, 1,
      0, -1),
    ncol = 2,
    byrow = TRUE
  )
  
  position <- c(0, 0)
  current_direction <- sample(1:4, 1)
  position <- position + directions[current_direction, ]
  if (n >= 2) {
    for (i in 2:n) {
      if (runif(1) < p) {
      } else {
        current_direction <- sample(1:4, 1)
      }
      position <- position + directions[current_direction, ]
    }
  }
  return(position)
}
#version 1 test
momentum_walk1(10, 0.6)
momentum_walk1(1000, 0.6)

#version2
momentum_walk2 <- function(n, p) {
  
  if (length(n) != 1 || !is.numeric(n) || n < 0 || n != floor(n)) {
    stop("n must be a non-negative integer")
  }
  if (length(p) != 1 || !is.numeric(p) || p < 0 || p > 1) {
    stop("p must be between 0 and 1")
  }
  directions <- matrix(
    c(1, 0,
      -1, 0,
      0, 1,
      0, -1),
    ncol = 2,
    byrow = TRUE
  )
  current_direction <- sample(1:4, 1)
  one_step <- function() {
    if (runif(1) >= p) {
      current_direction <<- sample(1:4, 1)
    }
    directions[current_direction, ]
  }
  position <- directions[current_direction, ]
  if (n >= 2) {
    steps <- replicate(n - 1, one_step())
    position <- position + rowSums(steps)
  }
  
  return(position)
}
#version2 test
momentum_walk2(10, 0.6)

momentum_walk2(1000, 0.6)
```

### Explanation problem1(b)

### Version 1

Version 1 uses a for loop. At each step, the function either keeps the previous direction with probability p or randomly selects a new direction with probability 1-p. The position is then updated by adding the corresponding direction vector.

### Version 2

Version 2 uses `replicate()` to repeat the step simulation. The previous direction is stored and updated after each simulated step, so the momentum dependence is preserved.

```{r}
#part c 
#set example
set.seed(123)
momentum_walk1(10, 0.6)
set.seed(123)
momentum_walk1(10, 0.6)
set.seed(123)
momentum_walk1(1000, 0.6)
set.seed(123)
momentum_walk1(1000, 0.6)
```

### Explanation problem1(c)

I used `set.seed(123)` before each simulation. This fixes the starting point of R's pseudorandom number generator, so the same sequence of random numbers is generated each time.

```{r}
#problem1 d
set.seed(123)
p_values <- c(0, 0.5, 0.9)
for (p in p_values) {
  position <- momentum_walk1(1000, p)
  distance <- sqrt(sum(position^2))
  cat("p =", p, 
      "position =", position,
      "distance =", distance, "\n")
}

```

### Explanation problem1(d)

Increasing momentum tends to increase the final distance from the origin.The larger momentum, the stronger movement persistence, the fewer cancellations, and bigger distance.

```{r}
p_values <- c(0, 0.25, 0.5, 0.75, 0.9)

mean_distance <- sapply(p_values, function(p) {
  
  distances <- replicate(1000, {
    
    position <- momentum_walk1(1000, p)
    
    sqrt(sum(position^2))
  })
  
  mean(distances)
})

results <- data.frame(
  p = p_values,
  mean_distance = mean_distance
)

print(results)

```

### Explanation problem1(e)

The mean final distance generally increases as momentum increases. When p is larger, the particle is more likely to continue moving in the same direction, so consecutive steps are more persistent and there is less cancellation between steps.

```{r}
library(microbenchmark)
#n=1000
library(microbenchmark)
benchmark_1000 <- microbenchmark(
  loop = momentum_walk1(1000, 0.6),
  replicate = momentum_walk2(1000, 0.6),
  times = 100
)
print(benchmark_1000)
#n=100000
benchmark_100000 <- microbenchmark(
  loop = momentum_walk1(100000, 0.6),
  replicate = momentum_walk2(100000, 0.6),
  times = 100
)
print(benchmark_100000)
```

### Explanation problem1(f)

Replicate implementation is faster.The absolute difference in execution time becomes much larger when n increases from 1,000 to 100,000.The loop implementation directly updates the particle's position at each step. The `replicate()` implementation repeatedly calls another function, stores the results of all steps, and then combines them using `rowSums()`. These additional function-call and result-storage operations introduce overhead, making the `replicate()` implementation slower.
