---
title: "hw2"
author: "Rao Yuner"
format: html
self-contained: true
code-fold: true
engine: knitr
---

```{r}
#problem1_a  momentum_walk(n,p)
momentum_walk <- function(n, p) {
  if (length(n) != 1 || !is.numeric(n) || n < 0 || n != floor(n)) {
    stop("n must be a non-negative integer")
  }
  if (length(p) != 1 || !is.numeric(p) || p < 0 || p > 1) {
    stop("p must be between 0 and 1")
  }
  directions <- matrix(
    c(1, 0,
      -1, 0,
      0, 1,
      0, -1),
    ncol = 2,
    byrow = TRUE
  )
  position <- c(0, 0)
  current_direction <- sample(1:4, 1)
  position <- position + directions[current_direction, ]
  if (n >= 2) {
    for (i in 2:n) {
      if (runif(1) < p) {
      } else {
        current_direction <- sample(1:4, 1)
      }
      position <- position + directions[current_direction, ]
    }
  }
  return(position)
}

#set example
momentum_walk(10, 0.6)
```

### Explanation problem1(a)

if we run the same parameters for several times, the outputs are different.

```{r}
#problem1_b
#version1
momentum_walk1 <- function(n, p) {
  if (length(n) != 1 || !is.numeric(n) || n < 0 || n != floor(n)) {
    stop("n must be a non-negative integer")
  }
  if (length(p) != 1 || !is.numeric(p) || p < 0 || p > 1) {
    stop("p must be between 0 and 1")
  }
  directions <- matrix(
    c(1, 0,
      -1, 0,
      0, 1,
      0, -1),
    ncol = 2,
    byrow = TRUE
  )
  position <- c(0, 0)
  current_direction <- sample(1:4, 1)
  position <- position + directions[current_direction, ]
  if (n >= 2) {
    for (i in 2:n) {
      if (runif(1) < p) {
      } else {
        current_direction <- sample(1:4, 1)
      }
      position <- position + directions[current_direction, ]
    }
  }
  return(position)
}
#version 1 test
momentum_walk1(10, 0.6)
momentum_walk1(1000, 0.6)
#version2
momentum_walk2 <- function(n, p) {
  if (length(n) != 1 || !is.numeric(n) || n < 0 || n != floor(n)) {
    stop("n must be a non-negative integer")
  }
  if (length(p) != 1 || !is.numeric(p) || p < 0 || p > 1) {
    stop("p must be between 0 and 1")
  }
  directions <- matrix(
    c(1, 0,
      -1, 0,
      0, 1,
      0, -1),
    ncol = 2,
    byrow = TRUE
  )
  current_direction <- sample(1:4, 1)
  one_step <- function() {
    if (runif(1) >= p) {
      current_direction <<- sample(1:4, 1)
    }
    directions[current_direction, ]
  }
  position <- directions[current_direction, ]
  if (n >= 2) {
    steps <- replicate(n - 1, one_step())
    position <- position + rowSums(steps)
  }
  return(position)
}
#version2 test
momentum_walk2(10, 0.6)
momentum_walk2(1000, 0.6)
```

### Explanation problem1(b)

### Version 1

Version 1 uses a for loop. At each step, the function either keeps the previous direction with probability p or randomly selects a new direction with probability 1-p. The position is then updated by adding the corresponding direction vector.

### Version 2

Version 2 uses replicate() to repeat the step simulation. The previous direction is stored and updated after each simulated step, so the momentum dependence is preserved.

```{r}
#problem1_c 
#set example
set.seed(123)
momentum_walk1(10, 0.6)
set.seed(123)
momentum_walk1(10, 0.6)
set.seed(123)
momentum_walk1(1000, 0.6)
set.seed(123)
momentum_walk1(1000, 0.6)
```

### Explanation problem1(c)

I used set.seed(123) before each simulation. This fixes the starting point of R's pseudorandom number generator, so the same sequence of random numbers is generated each time.

```{r}
#problem1_d
set.seed(123)
p_values <- c(0, 0.5, 0.9)
for (p in p_values) {
  position <- momentum_walk1(1000, p)
  distance <- sqrt(sum(position^2))
  cat("p =", p, 
      "position =", position,
      "distance =", distance, "\n")
}
```

### Explanation problem1(d)

Increasing momentum tends to increase the final distance from the origin.The larger momentum, the stronger movement persistence, the fewer cancellations, and bigger distance.

```{r}
#problem1_e
p_values <- c(0, 0.25, 0.5, 0.75, 0.9)
mean_distance <- sapply(p_values, function(p) {
  distances <- replicate(1000, {
    position <- momentum_walk1(1000, p)
    sqrt(sum(position^2))
  })
  mean(distances)
})
results <- data.frame(
  p = p_values,
  mean_distance = mean_distance
)
print(results)
```

### Explanation problem1(e)

The mean final distance generally increases as momentum increases. When p is larger, the particle is more likely to continue moving in the same direction, so consecutive steps are more persistent and there is less cancellation between steps.

```{r}
#problem1_f
library(microbenchmark)
#n=1000
library(microbenchmark)
benchmark_1000 <- microbenchmark(
  loop = momentum_walk1(1000, 0.6),
  replicate = momentum_walk2(1000, 0.6),
  times = 100
)
print(benchmark_1000)
#n=100000
benchmark_100000 <- microbenchmark(
  loop = momentum_walk1(100000, 0.6),
  replicate = momentum_walk2(100000, 0.6),
  times = 100
)
print(benchmark_100000)
```

### Explanation problem1(f)

Replicate implementation is faster.The absolute difference in execution time becomes much larger when n increases from 1,000 to 100,000.The loop implementation directly updates the particle's position at each step. The replicate() implementation repeatedly calls another function, stores the results of all steps, and then combines them using rowSums(). These additional function-call and result-storage operations introduce overhead, making the replicate() implementation slower.

```{r}
#problem2
set.seed(123)
# Number of simulated days
n_days <- 10000
# 6 AM to 8 AM: 3 hours
morning <- matrix(
  rpois(n_days * 3, lambda = 4),
  nrow = n_days,
  ncol = 3
)
# 9 AM: 1 hour
nine_am <- rnorm(
  n_days,
  mean = 45,
  sd = sqrt(16)
)
# 10 AM to 3 PM: 6 hours
midday <- matrix(
  rpois(n_days * 6, lambda = 15),
  nrow = n_days,
  ncol = 6
)
# 4 PM: 1 hour
four_pm <- rnorm(
  n_days,
  mean = 55,
  sd = sqrt(25)
)
# 5 PM to 10 PM: 6 hours
evening <- matrix(
  rpois(n_days * 6, lambda = 9),
  nrow = n_days,
  ncol = 6
)
# Total passengers for each day
daily_total <- rowSums(morning) +
  nine_am +
  rowSums(midday) +
  four_pm +
  rowSums(evening)
# Monte Carlo estimate of expected daily passengers
mean(daily_total)
```

### Explanation problem2

I simulated 10,000 days without using loops. For each time period, I generated the appropriate number of random observations using rpois( ) or rnorm( ). For periods containing multiple hours, I stored the observations in a matrix with one row per day and used rowSums( ). to calculate the passenger total for each day. Finally, I calculated the mean of the 10,000 daily totals to obtain the Monte Carlo estimate of the expected number of passengers per day.The theoretical expected value is 256 passengers per day, so the Monte Carlo estimate should be close to 256.

```{r}
#problem3_a
youtube <- read.csv(
  "https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2021/2021-03-02/youtube.csv"
)
dim(youtube)
names(youtube)
youtube_clean <- youtube[, c(
  "year",
  "funny",
  "show_product_quickly",
  "patriotic",
  "celebrity",
  "danger",
  "animals",
  "use_sex",
  "view_count",
  "like_count",
  "dislike_count",
  "favorite_count",
  "comment_count",
  "category_id"
)]
dim(youtube_clean)
```

### Explanation problem3(a)

After removing variables that could identify individual commercials, the resulting dataset has **247 observations and 14 variables**.

```{r}
#problem3_b
par(mfrow = c(2, 3))
hist(
  youtube_clean$view_count,
  main = "View Counts",
  xlab = "View Count"
)
hist(
  youtube_clean$like_count,
  main = "Like Counts",
  xlab = "Like Count"
)
hist(
  youtube_clean$dislike_count,
  main = "Dislike Counts",
  xlab = "Dislike Count"
)
hist(
  youtube_clean$favorite_count,
  main = "Favorite Counts",
  xlab = "Favorite Count"
)
hist(
  youtube_clean$comment_count,
  main = "Comment Counts",
  xlab = "Comment Count"
)
par(mfrow = c(1, 1))
summary(youtube_clean$favorite_count)
unique(youtube_clean$favorite_count)
youtube_clean$log_view_count <-
  log1p(youtube_clean$view_count)
youtube_clean$log_like_count <-
  log1p(youtube_clean$like_count)
youtube_clean$log_dislike_count <-
  log1p(youtube_clean$dislike_count)
youtube_clean$log_comment_count <-
  log1p(youtube_clean$comment_count)
par(mfrow = c(2, 2))
hist(
  youtube_clean$log_view_count,
  main = "Log View Counts",
  xlab = "log(View Count + 1)"
)
hist(
  youtube_clean$log_like_count,
  main = "Log Like Counts",
  xlab = "log(Like Count + 1)"
)
hist(
  youtube_clean$log_dislike_count,
  main = "Log Dislike Counts",
  xlab = "log(Dislike Count + 1)"
)
hist(
  youtube_clean$log_comment_count,
  main = "Log Comment Counts",
  xlab = "log(Comment Count + 1)"
)
par(mfrow = c(1, 1))
```

### Explanation problem3(b)

The view counts, like counts, dislike counts, and comment counts are strongly right-skewed. Therefore, I use a log transformation, specifically log(x+1), to reduce the skewness and allow zero counts. The favorite_count variable is not appropriate as an outcome because it has no variation, with all observations having the same value. Thus, the four appropriate outcomes are log-transformed view counts, like counts, dislike counts, and comment counts.

```{r}
#problem3_c
characteristics <- c(
  "funny",
  "show_product_quickly",
  "patriotic",
  "celebrity",
  "danger",
  "animals",
  "use_sex"
)
outcomes <- c(
  "log_view_count",
  "log_like_count",
  "log_dislike_count",
  "log_comment_count"
)
youtube_clean[characteristics] <-
  lapply(
    youtube_clean[characteristics],
    as.numeric
  )
results <- list()
for (outcome in outcomes) {
  for (characteristic in characteristics) {
    formula <- as.formula(
      paste(
        outcome,
        "~",
        characteristic,
        "+ year"
      )
    )
    model <- lm(
      formula,
      data = youtube_clean
    )
    coefficient <-
      coef(summary(model))[
        characteristic,
        "Estimate"
      ]
    p_value <-
      coef(summary(model))[
        characteristic,
        "Pr(>|t|)"
      ]
    results[[length(results) + 1]] <-
      data.frame(
        outcome = outcome,
        characteristic = characteristic,
        estimate = coefficient,
        p_value = p_value
      )
  }
}
results <- do.call(
  rbind,
  results
)
results$significant <-
  results$p_value < 0.05
results
significant_results <-
  results[
    results$p_value < 0.05,
  ]
significant_results
view_results <-
  results[
    results$outcome == "log_view_count",
  ]
view_results
best_characteristic <-
  view_results$characteristic[
    which.min(view_results$p_value)
  ]
best_characteristic
view_results[
  which.min(view_results$p_value),
]
```

### Explanation problem3(c)

I fitted 28 regression models, controlling for year in each model. At the 0.05 significance level, none of the commercial characteristics was statistically significant for the four outcomes. For log-transformed view counts, **danger** had the smallest p-value (p = 0.0941) and an estimated coefficient of 0.6668. The positive coefficient suggests a positive association with log view counts, but the association was not statistically significant.

```{r}
#problem3_d
# Remove observations with missing values
d <- youtube_clean[
  complete.cases(
    youtube_clean[
      ,
      c(
        "log_view_count",
        best_characteristic,
        "year"
      )
    ]
  ),
]
# Response vector
y <- d$log_view_count
# Design matrix
X <- cbind(
  1,
  d[[best_characteristic]],
  d$year
)
# OLS using matrix algebra
beta_hat <-
  solve(t(X) %*% X) %*%
  t(X) %*%
  y
# Display matrix algebra results
beta_hat
# Fit the same model using lm()
formula_d <- as.formula(
  paste(
    "log_view_count ~",
    best_characteristic,
    "+ year"
  )
)
model_d <- lm(
  formula_d,
  data = d
)
beta_lm <- coef(model_d)
# Display lm results
beta_lm
# Compare results
comparison <- cbind(
  matrix_algebra = as.vector(beta_hat),
  lm = as.numeric(beta_lm)
)
comparison
# Check whether they are equal
all.equal(
  as.vector(beta_hat),
  as.numeric(beta_lm),
  tolerance = 1e-10)
```

### Explanation problem3(d)

I used danger, which had the smallest p-value for predicting log-transformed view counts in Part (c). The design matrix included an intercept, danger, and year. Using the OLS matrix formula
β\^=(XTX)−1XTy,\\hat{\\beta}=(X\^TX)\^{-1}X\^Ty,
I obtained the coefficient estimates:
β\^0=−31.3915,β\^1=0.6668,β\^2=0.02066.\\hat{\\beta}\_0=-31.3915,\\qquad
\\hat{\\beta}\_1=0.6668,\\qquad
\\hat{\\beta}\_2=0.02066.
Therefore, the fitted model is
log⁡(view count+1)\^=−31.3915+0.6668(danger)+0.02066(year).\\widehat{\\log(\\text{view count}+1)}
=
-31.3915+0.6668(\\text{danger})+0.02066(\\text{year}).
The coefficients obtained using matrix algebra were exactly the same as those obtained using lm(). The all.equal() result was TRUE, confirming that the matrix formula produces the same OLS estimates as lm().

Holding year constant, commercials with the danger characteristic have an estimated 0.6668 higher log-transformed view count than commercials without this characteristic. Holding danger constant, a one-year increase in year is associated with an estimated 0.02066 increase in log-transformed view count.
